\thispagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent
%
Firstly, I would like to express my sincere gratitude to my supervisor Prof.\,Dr.\,Michael Gertz for the continuous support, patience, motivation, and immense knowledge. His guidance helped me in all the time of writing of this thesis.\\
Furthermore, I owe my deepest gratitude to Andreas Spitz, without his assistance and dedicated involvement in every step throughout the process, this thesis would have never been accomplished. His insightful comments, encouragement, and instructions guided me in every step of the way. \\
I am also grateful to all the members of the IWR committee for allowing me the long and extensive use of their computing servers.\\
I would like to extend my thanks to Felix Feldmann, for his understanding and patience throughout these stressful times and for offering a helping hand in any difficult situation.\\
I am thankful to all the people who read this thesis completely or in parts. Great thanks to Dr.\,Bastian Rieck, for teaching me the proper ways of scientific writing and for always encouraging me to be better. I wish to thank my friends, Stefan Radev and Karsten Hanser for reading through this thesis even if the number of pages seemed daunting. \\
Most importantly, none of this could have happened without my family. I would like to thank my mother, my aunt, Ziba Mir-hosseini and Richard Tapper for providing me with the opportunity to follow my dreams and for their continuous support. 
\begin{flushleft}
  Thank you.
\end{flushleft}

